---
title: "Hands-on Exercise 2"
---

# Beyond ggplot2 Fundamentals

## Installing and Loading the required library

### check the required packages: ggrepel, ggthemes, hrbrthemes, patchwork

```{r}
pacman::p_load(ggrepel, patchwork, ggthemes,
               hrbrthemes, tidyverse)
```

## Importing data

```{r}
exam_data <- read.csv("data/Exam_data.csv")
```

##### *reader is one of the tidyverse package*

## Beyond ggplot2 Annotation:ggrepel

```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5) +
  geom_label(aes(label = ID),
             hjust = 0.5,
             vjust = -.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

##### *orignal plot*

```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5) +
  geom_label_repel(aes(label = ID),
             hjust = 0.5,
             vjust = -.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

##### *using ggrepel*

## Beyond ggplot2 Themes

##### *eight built_in themes: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void()*

#### theme_gray

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

#### theme_bw

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_bw() +
  ggtitle("Distribution of Maths scores")
```

#### theme_classic

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_classic() +
  ggtitle("Distribution of Maths scores")
```

#### theme_dark

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_dark() +
  ggtitle("Distribution of Maths scores")
```

#### theme_light

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_light() +
  ggtitle("Distribution of Maths scores")
```

#### theme_linedraw

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_linedraw() +
  ggtitle("Distribution of Maths scores")
```

#### theme_minimal

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_minimal() +
  ggtitle("Distribution of Maths scores")
```

#### theme_void

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  theme_void() +
  ggtitle("Distribution of Maths scores")
```

### Working with ggtheme package

#### Economist theme

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25",
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### Working with hrbyhems package

#### hybrthmes (where lables are placed)

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

#### hybrthmes (add details)

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

## Beyond Single Graph

### p1

```{r}
p1 <- ggplot(data = exam_data,
             aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = "grey25",
                 fill = "grey90") +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

### p2

```{r}
p2 <- ggplot(data = exam_data,
             aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100,
                 color = "grey25", fill = "grey90") +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

### p3

```{r}
p3 <- ggplot(data = exam_data,
             aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              size = .5) +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### show

```{r}
p1
p2
p3
```
