---
title: "Take_home_Ex03"
author: "Zou Jiaxun"
date: "Jun 7, 2024"
data-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
format:
  html:
    code-fold: true
    code-summary: "Click to view code"
---

# Vast Challenge 2024 Mini-Challenge 3

In this challenge, I will attempt to do the following:

-   

    1.  Use visual analytics to identify the relationship of current fish companies

-   

    2.  Based on the visualization on Network, infer ownership or influence that a network

## 1 Getting Start

### 1.1 Package Preparation

```{r}
pacman::p_load(tidyverse, readtext, skimr, knitr,
               quanteda, tidytext, jsonlite, dplyr,
               tidyr, tidygraph, ggraph, igraph, lubridate,
               visNetwork, ggplot2, gganimate, gridExtra)
```

## 2 Data Wangling

### 2.1 Data Import

```{r}
mc3_data <- fromJSON("data/MC3/mc3 1.json")
class(mc3_data)
```

### 2.2 Data Check

::: panel-tabset
## edges

```{r}
mc3_edges <- as_tibble(mc3_data$link)
glimpse(mc3_edges)
```

## nodes

```{r}
mc3_nodes <- as_tibble(mc3_data$nodes)
glimpse(mc3_nodes)
```
:::

### 2.3 Data Filter

::: panel-tabset
## edges

```{r}
# Check all end data is N/A or in 2035 YEAR
all_end_dates_valid <- mc3_edges %>%
  mutate(end_date_year = year(as_date(end_date))) %>%
  summarise(all_valid = all(is.na(end_date) | end_date_year == 2035)) %>%
  pull(all_valid)

# Print the result
print(all_end_dates_valid)

# Convert "NA" strings to NA in end_date
mc3_edges_add <- mc3_edges %>%
  mutate(end_date = na_if(end_date, "NA")) %>%
  mutate(operation = as.character(ifelse(is.na(end_date), 0, 1)))

# Display the modified dataset
glimpse(mc3_edges_add)

```

```{r}
mc3_edges_filt <- mc3_edges_add %>% select(type, source, target, start_date, end_date, operation)
glimpse(mc3_edges_filt)
```

## nodes

```{r}
mc3_nodes_filt <- mc3_nodes %>% select(type, id, country, ProductServices)
glimpse(mc3_nodes_filt)
```
:::

### 2.4 Edges & Nodes Explore

::: panel-tabset
## edges

```{r}
skim(mc3_edges_filt)
```

```{r}
p1 <- ggplot(data = mc3_edges_filt, aes(x = type, fill = type)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
p1
```

## nodes

```{r}
skim(mc3_nodes_filt)
```

```{r}
p2 <- ggplot(data = mc3_nodes_filt, aes(x = type, fill = type)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
p2 
```
:::

::: callout-tip
According to the VAST2024 - MC3 Data Description, the person node include Entity.Person and Entity.Person.CEO, the other were included in the Company.
:::

## 3 MC3 Task(3)

Develop a visual approach to examine inferences. Infer how the influence of a company changes through time. Can you infer ownership and influence that a network may have?

::: callout-tip
Based on this plot, the edge types like {Shareholdership, BeneficialOwnership, WorksFor, FamilyRelationship} will be used, and the company name to define company; Meanwhile, the type in node like (Entity.Organization, Entity.Organization.Company, Entity.Organization.FishingCompany, Entity.Organization.LogisticsCompany, Entity.Organization.NewsCompany, Entity.Organization.FinancialCompany, Entity.Organization.NGO) will be used, source and target will be choosed to decide the "from" and "to".
:::

### 3.1 Relationship

::: panel-tabset
#### 3.1.1 Nodes

The code chunk below will be used to extract the nodes data.frame of mc3_data and save it as a tibble data.frame called mc3_nodes.

```{r}
# Clean and select columns from mc3_data$nodes
mc3_nodes_clean <- as_tibble(mc3_nodes_filt) %>%
  mutate(type = type,
         id = id) %>%
  select(id, type) %>%
  separate(type, into = c("prefix", "supertype", "subtype"), sep = "\\.", extra = "merge", fill = "right") %>%
  select(id, supertype, subtype)

# Display the first few rows using kable
kable(head(mc3_nodes_clean))
```

#### 3.1.2 Edges

The code chunk below will be used to extract the links data.frame of mc3_data and save it as a tibble data.frame called mc3_edges.

```{r}
mc3_edges_clean <- as_tibble(mc3_edges_filt) %>% 
  mutate(
    start_date = as_datetime(start_date),
    end_date = as_datetime(end_date)
    ) %>%
  group_by(source, target, type, operation, start_date, end_date) %>%
    summarise(weights = n(), .groups = 'drop') %>%
  filter(source != target) %>%
  ungroup()
kable(head(mc3_edges_clean))
```
:::

### 3.2 Visualising network graph

First, create a graph object using tbl_graph() function. Then calculate betweenness and closeness centrality scores.

```{r}
mc3_graph <- tbl_graph(nodes = mc3_nodes_clean,
                       edges = mc3_edges_clean,
                       directed = FALSE) %>%
  mutate(betweenness_centrality = centrality_betweenness(),
         closeness_centrality = centrality_closeness())
```

We will filter nodes with high betweenness centrality scores (\>8,000,000) and visualise them to see the relationships that they have.

```{r}
set.seed(1234)
mc3_graph %>%
  filter(betweenness_centrality >= 8000000) %>%
ggraph(layout = "fr") +
  geom_edge_link(aes(#width= weights,
                     alpha=0.5)) +
  geom_node_point(aes(
    size = betweenness_centrality,
    color = supertype,
    alpha = 0.3)) +
  geom_node_label(aes(label = id),repel=TRUE, size=2.5, alpha = 0.8) +
  scale_size_continuous(range=c(1,10)) +
  theme_graph() +
  labs(title = 'Initial network visualisation',
       subtitle = 'Entities with betweenness scores > 8,000,000')
```

Below is a dataframe showing us the top 10 entities with the highest betweenness scores.

```{r}
mc3_graph %>%  activate(nodes) %>%  as_tibble() %>% arrange(desc(betweenness_centrality)) %>% slice(1:10) %>% kable() 
```

The top 10 betweenness entities above include 3 persons and 7 companies. In the next section, we will filter entities into only organization entities. We may revisit the person entities later when we have specific targets/companies to investigate.

### 3.3 Check the Illegal event timeline

(1) before 2033-10-29
(2) between 2033-10-29 and 2035-05-25
(3) after 2035-05-25

```{r}
# Time bins
bin1 <- mc3_edges_clean %>%
  filter(as.Date(start_date) < as.Date("2033-10-29"))

bin2 <- mc3_edges_clean %>%
  filter(as.Date(start_date) >= as.Date("2033-10-29") & as.Date(start_date) <= as.Date("2035-05-25"))

bin3 <- mc3_edges_clean %>%
  filter(as.Date(start_date) > as.Date("2035-05-25"))

# Filter edges to include any that contain "SouthSeafood"
southseafood_edges_bin1 <- bin1 %>%
  filter(str_detect(source, "SouthSeafood") | str_detect(target, "SouthSeafood"))

southseafood_edges_bin2 <- bin2 %>%
  filter(str_detect(source, "SouthSeafood") | str_detect(target, "SouthSeafood"))

southseafood_edges_bin3 <- bin3 %>%
  filter(str_detect(source, "SouthSeafood") | str_detect(target, "SouthSeafood"))

# Glimpse the results
glimpse(southseafood_edges_bin1)
glimpse(southseafood_edges_bin2)
glimpse(southseafood_edges_bin3)
```

### 3.4 Exploring the VisNetwork for company influence based on timeline

#### Filter nodes to include only Organizations

```{r}
# Filter nodes to include only Organizations or Persons
mc3_nodes_select <- mc3_nodes_clean %>%
  filter(grepl("Organization|Person", supertype, ignore.case = TRUE))
```

#### 3.3.2 Extract edges related to nodes

```{r}
mc3_edges_select <- mc3_edges_clean %>%
  filter(source %in% mc3_nodes_select$id | target %in% mc3_nodes_select$id) %>%
  distinct() %>%
  rename(from = source,
         to = target)
```

#### 3.3.3 Extract edges with high connection

```{r}
mc3_edges_select_high <- mc3_edges_select %>%
  group_by(from) %>%
  mutate(count = n()) %>%
  filter(count >= 5) %>%
  ungroup()
```

#### 3.3.4 Distinct source and target

```{r}
mc3_nodes_source <- mc3_edges_select_high %>%
  distinct(from) %>%
  rename(id = from)

mc3_nodes_target <- mc3_edges_select_high %>%
  distinct(to) %>%
  rename(id = to)
```

#### 3.3.5 Build visNetwork

```{r}
# Combine and deduplicate nodes
mc3_nodes_combined <- bind_rows(mc3_nodes_source, mc3_nodes_target) %>%
  distinct(id) %>%
  left_join(mc3_nodes_clean, by = "id")
  
# Assign group based on supertype
mc3_nodes_combined$group <- ifelse(mc3_nodes_combined$supertype == "Organization", "Organization", "Person")

# Visualize with visNetwork
visNetwork(mc3_nodes_combined, mc3_edges_select_high) %>%
  visPhysics(solver = "forceAtlas2Based",
             forceAtlas2Based = list(gravitationalConstant = -100)) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visGroups(groupname = "Organization", color = "yellow") %>%
  visGroups(groupname = "Person", color = "grey") %>%
  visLegend() %>%
  visEdges(arrows = "to", label = "start_date") %>%
  visOptions(
    highlightNearest = list(enabled = TRUE, degree = 2, hover = TRUE),
    nodesIdSelection = TRUE,
    selectedBy = "group",
    collapse = TRUE
  ) %>%
  visInteraction(dragNodes = TRUE, dragView = TRUE, zoomView = TRUE) %>%
  visEvents(selectNode = "function(properties) {
    var selectedNodeId = properties.nodes[0];
    this.body.data.nodes.update({id: selectedNodeId, color: {background: 'red', border: 'black'}});
  }")
```

It is interesting to find that there is one complicated center which connected entities like "Hart Ltd", "Johnson, Morales and Castro", which showed a high cooperation and organiztion between current fish companies. And it might be interested to dig the development based on timeline or the financial situation to define its influence. What's more, even if some points are on the very outside, they still have a sound connection with certain company maybe the supplier or retailer.

## 4. Expand for SouthSeafood Express Corp

::: callout-idea
To ensure the network associated with SouthSeafood Express Corp and visualize how this network and competing businesses change as a result of their illegal fishing behavior, an expand in timeline of visNetwork of the SouthSeafood is required.
:::






## 5.Conclusion

Exploring the networks between the various type of nodes or players in the space has been useful to visualising the relationships between the different parties. It has yielded interesting insights on how certain companies may influence the around companies.

For future work, the additional column of timeline and financial situation can be used to provide an additional layer of to the overall visualisation of networks and information in this project.
