---
title: "In-class_Ex05"
author: "Zou Jiaxun"
date: "May 11, 2024"
data-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true

---
Put all data into one tibular dataframe. 

```{r}
pacman::p_load(tidyverse, readtext,
               quanteda, tidytext)
```

```{r}
data_folder <- "data/MC1/articles"
```
Text sensing to extract text


```{r}
text_data <- readtext(paste0("data/MC1/articles",
                "/*"))
```
OR


```{r}
text_data <- readtext("data/MC1/articles")
```

Basic tokenisation

```{r}
usenet_words <- text_data %>%
  unnest_tokens(word, text) %>%  #reading the text data
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word) #remove stop words
```

```{r}
usenet_words %>%
  count(word, sort = TRUE)
```

Observations- Most common words are: fishing, sustainable, company

```{r}
temp_table <- usenet_words %>%
  count(word, sort = TRUE)
```

Creating a table to observe word counts

```{r}
corpus_text <- corpus(text_data)
summary(corpus_text, 5)
```

To separate the data; with 2 columns X & Y.  Some text are "1" hence the split does not occur

```{r}
text_data_splitted <- text_data %>%
  separate_wider_delim("doc_id",
                       delim = "__0__",
                       names = c("X", "Y"),
                       too_few = "align_end")
```




```{r}
pacman::p_load(jsonlite, tidyverse)
```


```{r}
##pacman::p_load(jsonlite, tidygraph,
##ggraph, tidyverse, readtext,
               ##quanteda, tidytext)
```



In the code chunk below, fromJSON() of jsonlite package is used to import *MC3.json* into R environment.

```{r}
mc1_data <- fromJSON("data/MC1/mc1.json")
```


mc2_data <- fromJSON("data/MC2/mc2.json")

mc3_data <- fromJSON("data/MC3/mc3.json")
